---
title: "HESA data"
format: revealjs
---


```{r}
#| echo: false

hesa_stats <- readr::read_csv(
  "./data/sb265_figure_13.csv",
  skip = 16,
  col_names = c("Subject Area", "2019_Female", "2019_Male", "2019_Other", "2019_Total", "2020_Female", "2020_Male", "2020_Other", "2020_Unknown", "2020_Total", "2021_Female", "2021_Male", "2021_Other", "2021_Unknown", "2021_Total")) |> 
  dplyr::filter(!grepl("Total", `Subject Area`)) |> 
  tidyr::pivot_longer(
    cols = `2019_Female`:`2021_Total`,
    names_to = c("Year", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  dplyr::select(-Unknown) |> 
  dplyr::mutate(
    `Subject Area` = gsub("[0-9]{2}\\s", "", `Subject Area`),
    `Proportion female` = Female/Total
  ) |> 
  dplyr::arrange(desc(`Proportion female`)) |> 
  dplyr::group_by(Year) |> 
  dplyr::mutate(Rank = dense_rank(desc(`Proportion female`))) |> 
  select(`Subject Area`, Year, Female, `Proportion female`, Rank) |> 
  rename(
    N = Female,
    Proportion = `Proportion female`
  ) |> 
  ungroup()
```

## HESA student data 2019-20^[Data from [www.hesa.ac.uk/data-and-analysis/students/what-study](https://www.hesa.ac.uk/data-and-analysis/students/what-study)]

::: {.panel-tabset}

## Proportion of women

```{r}
#| echo: false

hesa_stats <- hesa_stats |> 
  mutate(
    Year = factor(ordered(Year, levels = c(2019, 2020, 2021))),
    n_1000 = N/1000
  )

hesa_stats_filter <- hesa_stats |> 
  filter(Rank < 6 | Rank > 17) 

ggplot(data = hesa_stats_filter, aes(x = Year, y = Proportion, colour = `Subject Area`, label = `Subject Area`)) +
  geom_point(size = 3) +
  geom_line(aes(group = `Subject Area`), linewidth = 0.75) +
  geom_text_repel(
    data          = subset(hesa_stats_filter, Year == max(Year)),
    nudge_x       = 0.1,
    segment.size  = 0.2,
    segment.color = "grey50",
    direction     = "y",
    hjust         = 0
  ) +
  coord_cartesian(ylim = c(0.15, 0.85), xlim = c(1, 4)) +
  scale_y_continuous(breaks = seq(0.15, 0.85, 0.05)) +
  scale_color_viridis_d(option = "turbo") +
  labs(y = "Proportion of applicants (women)") +
  theme_minimal() +
  theme(legend.position = "none")
```

## Number of women

```{r}
#| echo: false

ggplot(data = hesa_stats_filter, aes(x = Year, y = n_1000, colour = `Subject Area`, label = `Subject Area`)) +
  geom_point(size = 3) +
  geom_line(aes(group = `Subject Area`), linewidth = 0.75) +
  geom_text_repel(
    data          = subset(hesa_stats_filter, Year == max(Year)),
    nudge_x       = 0.1,
    segment.size  = 0.2,
    segment.color = "grey50",
    direction     = "y",
    hjust         = 0
  ) +
  coord_cartesian(xlim = c(1, 4)) +
  scale_y_continuous(breaks = c(10, 50, 100, 200, 300)) +
  scale_color_viridis_d(option = "turbo") +
  labs(y = "Number of women applicants (thousands)") +
  theme_minimal() +
  theme(legend.position = "none")
```

:::


::: {.notes}

It seems like computing has an issue with both attracting female school-leavers, and retaining women that choose computing as a career

Needless to say there's a lot of research into why these problems exist.

I'm going to focus on one particular piece of work

:::
